---
title: "Homework 1 Submission 1"
format: pdf
author: Tammy Dang
execute:
    echo: false
---

```{r}
#load libraries
library(tidyverse)

#get working directory
getwd()
setwd('/Users/tammydang/Downloads/EMORY/ECON-470')
```
# Question 1: 
Provide a table of the count of plans under each plan type. Your table should look something like Table 1. (2 points)

```{r}
#CODE CHUNK
setwd('/Users/tammydang/Downloads/EMORY/ECON-470')
# read enroll ------------------------------
read_enroll <- function(path) {
    read_csv(
      path,
      skip = 1,
      col_names = c("contractid","planid","ssa","fips","state","county","enrollment"),
      col_types = cols(
        contractid = col_character(),
        planid     = col_double(),
        ssa        = col_double(),
        fips       = col_double(),
        state      = col_character(),
        county     = col_character(),
        enrollment = col_double()
      ),
      na = "*",
      show_col_types = FALSE,
      progress = FALSE
    )
  }

#function to read contract data
  read_contract <- function(path) {
    read_csv(
      path,
      skip = 1,
      col_names = c(
        "contractid","planid","org_type","plan_type","partd","snp","eghp",
        "org_name","org_marketing_name","plan_name","parent_org","contract_date"
      ),
      col_types = cols(
        contractid = col_character(),
        planid     = col_double(),
        org_type   = col_character(),
        plan_type  = col_character(),
        partd      = col_character(),
        snp        = col_character(),
        eghp       = col_character(),
        org_name   = col_character(),
        org_marketing_name = col_character(),
        plan_name  = col_character(),
        parent_org = col_character(),
        contract_date = col_character()
      ),
      show_col_types = FALSE,
      progress = FALSE
    )
  }

#load one month --------------------------------------------------
 load_month <- function(m, y) {
    c_path <- paste0("data/ma-data/ma-data/ma/enrollment/Extracted Data/CPSC_Contract_Info_", y, "_", m, ".csv")
    e_path <- paste0("data/ma-data/ma-data/ma/enrollment/Extracted Data/CPSC_Enrollment_Info_", y, "_", m, ".csv")

    contract.info <- read_contract(c_path) %>%
      distinct(contractid, planid, .keep_all = TRUE)   

    enroll.info <- read_enroll(e_path)

    contract.info %>%
      left_join(enroll.info, by = c("contractid","planid")) %>%
      mutate(month = as.integer(m), year = y)
  }

 # Read all months, then tidy into a year-----------------------------
 
 #load_year is for plan.year for the monthlist
 y = 2018
monthlist <- if (y == 2006) sprintf("%02d", 7:12) else sprintf("%02d", 1:12)

# ALL MONTHS STACKED ON TOP OF EACH OTHER
plan.year <- map_dfr(monthlist, ~ load_month(.x, y)) %>%
    arrange(contractid, planid, state, county, month) %>%
    group_by(state, county) %>%
    fill(fips, .direction = "downup") %>%                
    ungroup() %>%
    group_by(contractid, planid) %>%
    fill(plan_type, partd, snp, eghp, plan_name, .direction = "downup") %>%
    ungroup() %>%
    group_by(contractid) %>%
    fill(org_type, org_name, org_marketing_name, parent_org, .direction = "downup") %>%
    ungroup()
 
 # Read service area data -----------------------------------------
  read_service_area <- function(path) {
    read_csv(
      path, skip = 1,
      col_names = c(
        "contractid","org_name","org_type","plan_type","partial","eghp",
        "ssa","fips","county","state","notes"
      ),
      col_types = cols(
        contractid = col_character(),
        org_name   = col_character(),
        org_type   = col_character(),
        plan_type  = col_character(),
        partial    = col_logical(),
        eghp       = col_character(),
        ssa        = col_double(),
        fips       = col_double(),
        county     = col_character(),
        state      = col_character(),
        notes      = col_character()
      ),
      na = "*",
      show_col_types = FALSE,
      progress = FALSE
    )
  }

 # One-month loader --------------------------------------------------------
  load_month_sa <- function(m, y) {
    path <- paste0("data/ma-data/ma-data/ma/service-area/Extracted Data/MA_Cnty_SA_",y, "_", m, ".csv")
    
    read_service_area(path) %>%
      mutate(month = as.integer(m), year = y)
  }

  # Read all months, then tidy once ----------------------------------------
  service.year <- map_dfr(monthlist, ~ load_month_sa(.x, y))

  # Ensure stable order before fills
  service.year <- service.year %>%
    arrange(contractid, fips, state, county, month)

  # Fill missing identifiers/labels
  service.year <- service.year %>%
    group_by(state, county) %>%
    fill(fips, .direction = "downup") %>%
    ungroup() %>%
    group_by(contractid) %>%
    fill(plan_type, partial, eghp, org_type, org_name, .direction = "downup") %>%
    ungroup()

# MERGE DATASET ------------

```
```{r}
table(plan.year$plan_type)
```

# Question 2:
Remove all special needs plans (SNP), employer group plans (eghp), and all “800-series” plans. Provide an updated version of Table 1 after making these exclusions. (2 points)
```{r}
plan.year_q2 <- plan.year %>% filter(
                                !str_detect(
                                    plan_type,
                                    regex(
                                        "SNP|Medicare-Medicaid|Employer|Union|18\\d{2}",
                                        ignore_case = TRUE
                                        )
                                )
)
table(plan.year_q2$plan_type)

```

# Question 3:
Now repeat the same filters from part 2 but also focus only on counties in which plans are approved as per the service area files. With these data, provide a table of the average enrollments for each plan type. (2 points)
```{r}
# average enrollment by plan type (for each plan type)

#service_thing %>% 
#    group_by(plan_type) %>%
#        summarise(avg_enroll = mean(enrollment, na/rm=TRUE),
#        n_plans = n()
#                 )
```


```{r}
#tinytex::uninstall_tinytex()
#tinytex::install_tinytex()
```